---
title: "BMT Data Analysis"
author: "Dhiraj Khanna"
date: "27 September 2016"
output:
  md_document:
    variant: markdown_github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This analysis is with the latest data received today (29 Sep 16).

##Analysis on Death Date

### Overall

```{r,warning=FALSE, message=FALSE}
library(dplyr)
library(ggplot2)
library(survival)
library(lubridate)
library(tidyr)
library(broom)
library(survminer)

#Read in the data
BMT <- read.csv("BMT_29Sep.csv",na.strings=c("", "NA"))

#Select variables of interest
BMT1 <- select(BMT, TypeOfMM, Transplant.Date, C.Status,Death.date, Last.Follow.up,M.protein..,RenalFailure,Plasmacytoma,RelapseDate,ISS,TotalNoTherapy,RemissionStatusPreTx,RemissionStatusPostTx)
BMT1$Transplant.Date <- dmy(BMT1$Transplant.Date)
BMT1$Last.Follow.up <- dmy(BMT1$Last.Follow.up)
BMT1$Death.date <- dmy(BMT1$Death.date)
BMT1$RelapseDate <- dmy(BMT1$RelapseDate)

#Create censoring indicator variable for death
cenin_dead <- rep(1, nrow(BMT1))
for (i in 1:nrow(BMT1)) ifelse(BMT1$C.Status[i]!='Dead',cenin_dead[i] <- 0,cenin_dead[i] <- 1)

#Replace 'NA' death dates with LFU dates
index <- which(is.na(BMT1$Death.date))
BMT1$Death.date[index] <- BMT1$Last.Follow.up[index]

#Find time difference
BMT1 <- mutate(BMT1, TD=as.numeric(Death.date-Transplant.Date))
surv_data <- cbind(BMT1,cenin_dead)

```

Now let's plot the KM survival distribution for overall death.

```{r}
fit_overall <- survfit(Surv(surv_data$TD,surv_data$cenin_dead)~1,conf.type = "log-log")
ggsurvplot(fit_overall,risk.table = T, risk.table.height = 0.4,pval = T)
```

We can have a look at the model.
```{r}
summary(fit_overall)
print(fit_overall)
```

### Renal Failure

```{r}
RF <- na.omit(select(surv_data, RenalFailure, TD, cenin_dead))
fit_RF <- survfit(Surv(TD,cenin_dead)~RenalFailure,data=RF)
ggsurvplot(fit_RF,risk.table = T,risk.table.height = 0.4,pval = T)
summary(fit_RF)
print(fit_RF)
```

### Plasmacytoma

```{r}
PC <- na.omit(select(surv_data, Plasmacytoma, TD, cenin_dead))
fit_PC <- survfit(Surv(TD,cenin_dead)~Plasmacytoma,data=PC)
ggsurvplot(fit_PC,risk.table = T, risk.table.height = 0.4,pval = T)
summary(fit_PC)
print(fit_PC)
```

### Remission Status Before

```{r}
RSPre <- na.omit(select(surv_data, RemissionStatusPreTx, TD, cenin_dead))
fit_RSPre <- survfit(Surv(TD,cenin_dead)~RemissionStatusPreTx,data=RSPre)
ggsurvplot(fit_RSPre,risk.table = T,risk.table.height = 0.4,pval = T)
summary(fit_RSPre)
print(fit_RSPre)
```

### Remission Status After

```{r}
RSPost <- na.omit(select(surv_data, RemissionStatusPostTx, TD, cenin_dead))
fit_RSPost <- survfit(Surv(TD,cenin_dead)~RemissionStatusPostTx,data=RSPost)
ggsurvplot(fit_RSPost,risk.table = T, risk.table.height = 0.4,pval = T)
summary(fit_RSPost)
print(fit_RSPost)
```

### Number of Therapies

```{r}
thp <- na.omit(select(surv_data, TotalNoTherapy, TD, cenin_dead))
fit_thp <- survfit(Surv(TD,cenin_dead)~TotalNoTherapy,data=thp)
ggsurvplot(fit_thp,risk.table = T, risk.table.height = 0.4,pval = T)
summary(fit_thp)
print(fit_thp)
```

### ISS

Values 'I' and 'II' have been coerced to a single value - 'I'.

```{r}
surv_data$ISS[surv_data$ISS=="II"] <- "I"
iss <- na.omit(select(surv_data, ISS, TD, cenin_dead))

iss$ISS <- as.factor(iss$ISS)
fit_iss <- survfit(Surv(TD,cenin_dead)~ISS,data=iss)
ggsurvplot(fit_iss,risk.table = T, risk.table.height = 0.4,pval = T)
summary(fit_thp)
print(fit_thp)
```

## Analysis on Relapse Date

```{r}
#Create censoring indicator variable for relapse date
relapse <- rep(1, nrow(surv_data))
for (i in 1:nrow(surv_data)) ifelse(is.na(surv_data$RelapseDate[i]),relapse[i] <- 0,relapse[i] <- 1)

#Replace 'NA' death dates with LFU dates
index <- which(is.na(surv_data$RelapseDate))
surv_data$RelapseDate[index] <- surv_data$Last.Follow.up[index]

#time difference for transplant - relapse
surv_data <- mutate(surv_data, TD_R=as.numeric(RelapseDate-Transplant.Date))
rlp_data <- cbind(surv_data,relapse)
```

Now that the data is ready, we can do a similar analysis. Start off with an overall analysis.

```{r}
fit_overall <- survfit(Surv(rlp_data$TD_R,rlp_data$relapse)~1,conf.type = "log-log")
ggsurvplot(fit_overall,risk.table = T, risk.table.height = 0.4,pval = T)
```

We can have a look at the model.
```{r}
summary(fit_overall)
print(fit_overall)
```

### Renal Failure

```{r}
RF <- na.omit(select(rlp_data, RenalFailure, TD_R, relapse))
fit_RF <- survfit(Surv(TD_R,relapse)~RenalFailure,data=RF)
ggsurvplot(fit_RF,risk.table = T,risk.table.height = 0.4,pval = T)
summary(fit_RF)
print(fit_RF)
```

### Plasmacytoma

```{r}
PC <- na.omit(select(rlp_data, Plasmacytoma, TD_R, relapse))
fit_PC <- survfit(Surv(TD_R,relapse)~Plasmacytoma,data=PC)
ggsurvplot(fit_PC,risk.table = T, risk.table.height = 0.4,pval = T)
summary(fit_PC)
print(fit_PC)
```

### Remission Status Before

```{r}
RSPre <- na.omit(select(rlp_data, RemissionStatusPreTx, TD_R, relapse))
fit_RSPre <- survfit(Surv(TD_R,relapse)~RemissionStatusPreTx,data=RSPre)
ggsurvplot(fit_RSPre,risk.table = T, risk.table.height = 0.4,pval = T)
summary(fit_RSPre)
print(fit_RSPre)
```

### Remission Status After

```{r}
RSPost <- na.omit(select(rlp_data, RemissionStatusPostTx, TD_R, relapse))
fit_RSPost <- survfit(Surv(TD_R,relapse)~RemissionStatusPostTx,data=RSPost)
ggsurvplot(fit_RSPre,risk.table = T, risk.table.height = 0.4,pval = T)
summary(fit_RSPost)
print(fit_RSPost)
```

### Number of Therapies

```{r}
thp <- na.omit(select(rlp_data, TotalNoTherapy, TD_R, relapse))
fit_thp <- survfit(Surv(TD_R,relapse)~TotalNoTherapy,data=thp)
ggsurvplot(fit_thp,risk.table = T, risk.table.height = 0.4,pval = T)
summary(fit_thp)
print(fit_thp)
```

### ISS

Values 'II' and 'III' have been coerced to a single value - 'II'.

```{r}
iss <- na.omit(select(rlp_data, ISS, TD_R, relapse))
iss[iss=='II',] <- "I"
iss$ISS <- factor(iss$ISS)
iss$TD_R <- as.numeric(iss$TD_R)
iss$relapse <- as.numeric(iss$relapse)
fit_iss <- survfit(Surv(TD_R,relapse)~ISS,data=iss)
ggsurvplot(fit_iss,risk.table = T, risk.table.height = 0.4,pval = T)
summary(fit_thp)
print(fit_thp)
```

### Year of Transplant

We will make three groups for year of transplant - 1999 to 2005, 2006-2010, 2011 to present.

```{r}
surv_data <- mutate(surv_data, trans_year=ifelse(year(surv_data$Transplant.Date)<=2005,"Group A",
                     ifelse(year(surv_data$Transplant.Date)<=2010,"Group B","Group C")))
```

Now let's see the survival plot for both death date and relapse date.

### Year of Transplant - Death Date
```{r}
YOT <- na.omit(select(surv_data, trans_year, TD, cenin_dead))
fit_YOT <- survfit(Surv(TD,cenin_dead)~trans_year,data=YOT)
ggsurvplot(fit_YOT,risk.table = T,risk.table.height = 0.4,pval = T)
summary(fit_YOT)
print(fit_YOT)
```

### Year of Transplant - Relapse Date

```{r}
rlp_data <- cbind(surv_data,relapse)
YOT_R <- na.omit(select(rlp_data, trans_year, TD_R, relapse))
fit_YOT_R <- survfit(Surv(TD_R,relapse)~trans_year,data=YOT_R)
ggsurvplot(fit_YOT_R,risk.table = T, risk.table.height = 0.4,pval = T)
summary(fit_YOT_R)
print(fit_YOT_R)
```

